Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-02-28 article-models-tt.tex
rewritten part of intro

\section{Type theory before type constructors}\label{tt}
In this section we give a description of type theory before basic
constructors and with explicit extension, weakening and substitution. We will
formulate the type theory in such a way that contexts aren't lists of variable
declarations. The reasons for this are twofold: (1) in the models of type theory
such as a univalent universe or the graph model, contexts are types and graphs
respectively -- they are objects rather than lists of variable declarations; we
intend to give a general algebraic description of higher categories. Besides,
we will not have to be burdened with comments about variables being bounded or not, or fresh or
not occuring at all.

@@

\begin{lem}
When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
a type $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
precisely, the following inference rules are valid:
\begin{align*}
& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
% \\
%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
\end{align*}
\end{lem}

>>>>

\section{Type theory before type constructors}\label{tt}
In this section we give a description of type theory before basic
constructors and with explicit extension, weakening and substitution. We will
formulate the type theory in such a way that contexts aren't lists of variable
declarations. The main reason is that we don't see variable declarations in the
models either. This way we also set out to a more algebraic approach of type
theory and higher category theory. Thirdly, we will not have to be burdened with
comments about variables being bounded or not, or fresh or not occuring at all.

@@

\begin{lem}
When we weaken a term $\ctxext{y}{v}$ of $\ctxext{B}{Q}$ in context $\Gamma$ by
a family $A$, the term that we get is $\ctxext{\ctxwk{A}{y}}{\ctxwk{A}{v}}$. More
precisely, the following inference rules are valid:
\begin{align*}
& \inference{\jterm{\ctxext{\Gamma}{B}}{Q}{g}\quad\jterm{\ctxext{\Gamma}{B}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\ctxext{Q}{R}}}{\ctxwk{A}{\ctxext{g}{t}}}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{t}}}}
% \\
%& \inference{\jterm{\Gamma}{\ctxwk{A}{\ctxext{B}{Q}}}{\ctxwk{A}{w}}}{\jtermeq{\ctxext{\Gamma}{A}}{\ctxwk{A}{B}}{\cprojfst{\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A}{w}}}{\ctxwk{A}{(\cprojfst{B}{Q}{w})}}}\\
\end{align*}
\end{lem}
